---
title: "p8105_hw2_jdr2191"
author: "Jess Rosenberg"
due date: "09OCT2021"
output: github_document
---

## Problem 1

Load the tidyverse and readxl
```{r loadlibraries, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
```

Import excel file
```{r import_mrtrashwheel}
mrtrashwheel_df = read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel",
                             range = cell_cols("A:N")) %>%
                  janitor::clean_names() %>%
                  filter(dumpster != "NA") %>%
                  mutate(sports_balls = round(sports_balls, 0))

view(mrtrashwheel_df)
```

Read and clean the precipitation data for 2018 and 2019.
```{r precipitationdata_18}
precipitation18_df = read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "2018 Precipitation",
                     range = "A2:B14") %>%
                     janitor::clean_names() %>%
                     mutate(year = "2018")

view(precipitation18_df)

precipitation19_df = read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "2019 Precipitation",
                     range = "A2:B14") %>%
                     janitor::clean_names() %>%
                     na.omit %>%
                     mutate(year = "2019") 

view(precipitation19_df)
```

Combine preciptation datasets for 2018 and 2019.
```{r combined_precipitation}
precipitation_df = bind_rows(precipitation18_df, precipitation19_df) %>%
                   janitor::clean_names() %>%
                   mutate(month = month.name[as.numeric(month)])

view(precipitation_df)
```

The Mr. Trash Wheel dataset contains `r count(mrtrashwheel_df)` observations. In 2019, the median number of sports balls in a dumpster was  `r mrtrashwheel_df %>% filter(year == 2019) %>% summarise(median(sports_balls))` balls. The Mr. Trash Wheel dataset describes the types of trash Mr. Trash Wheel collects, including, `plastic bottles, polystyrene, cigarette butts, glass bottles, grocery bags, chip bags, and sports balls`. The dataset also provides information about the `date` the trash was collected, as well as its `weight` and `volume`.

The precipitation dataset contains `r count(precipitation_df)` observations. In 2018, the total precipitation was `r precipitation_df %>% filter(year == 2018) %>% summarise(sum(total))` inches. The precipitation dataset includes a `month` and `year` variable as well as a total `precipitation` variable.

## Problem 2

Import and clean the data in pols-month.csv.
```{r import_pols-month}
pols_month_df = read_csv("pols-month.csv") %>%
                janitor::clean_names() %>%
                separate(mon, sep = "-", into = c("year","month","day"), convert = TRUE) %>%
                mutate(month = month.name[as.numeric(month)]) %>%
                mutate(president = factor(prez_dem, labels = c("gop","dem"))) %>%
                subset(select = -c(prez_dem, prez_gop, day))


view(pols_month_df)
```

Import and clean the data in snp.csv.
```{r import_snp}
snp_df = read_csv("snp.csv") %>%
         janitor::clean_names() %>%
         separate(date, sep = "/", into = c("month","day","year"), convert = TRUE) %>%
         mutate(year = ifelse(year >= 50, year + 1900, year + 2000)) %>%
         mutate(month = month.name[as.numeric(month)]) %>%
         subset(select = -c(day)) %>%
         relocate(year, before = NULL) 

view(snp_df)
```

Import and Clean the Unemployment Data.
```{r import_unemployment}
unemployment_df = read_csv("unemployment.csv")

view(unemployment_df)
```

```{r tidy_unemployment}
unemployment_tidy_df = 
             unemployment_df %>%
             pivot_longer(Jan:Dec,
                          names_to = "month",
                          values_to = "unemployment") %>%
            mutate(month = match(month, month.abb)) %>%
            mutate(month = month.name[as.numeric(month)]) %>%
            janitor::clean_names()

view(unemployment_tidy_df)

```

Merge snp into pols.
```{r merge_snp_pols}
snp_pols_df =
  left_join(pols_month_df, snp_df, by = c("month","year"))

view(snp_pols_df)
```

Merge unemployment_tidy_df into snp_pols_df
```{r merge_unemployment_snp_pols}
snp_pols_unemployment_df =
  left_join(snp_pols_df, unemployment_tidy_df, by = c("month","year"))

view(unemployment_tidy_df)
view(snp_pols_unemployment_df)
```

The pols_month dataset contains year and month of the count, a variable indicating whether the President was a Democrat or a Republican on the associated date, variables indicating the number of Republican and Democratic Governors on the associated date, variables indicating the number of Republican and Democratic Senators on the associated date, and variables indicating the number of Republican and Democratic Representatives on the associated date.The dates in this dataset ranged from `r min(pull(pols_month_df, year))` to `r max(pull(pols_month_df, year))`. 

The snp dataset contained the year and month of the observation as well as a variabel indicating the closing values of the S&P stock index on the associated date. The dates in this dataset ranged from `r min(pull(snp_df, year))` to `r max(pull(snp_df, year))`.

The unemployment (tidied) dataset contained year and month of the observation as well as the percent of unemployment on the associated date. The dates in this dataset ranged from `r min(pull(unemployment_tidy_df, year))` to `r max(pull(unemployment_tidy_df, year))`.

We merged the pols_month dataset, the snp dataset and the unemployment (tidied) dataset to produce a resulting dataset entitled snp_pols_unemployment. This resulting dataset has `r ncol(snp_pols_unemployment_df)` columns and `r nrow(snp_pols_unemployment_df)` rows. In total, this dataset contains `r count(snp_pols_unemployment_df)` observations. This combined dataset has dates ranging from `r min(pull(snp_pols_unemployment_df, year))` to `r max(pull(snp_pols_unemployment_df, year))`. Some key variables in this dataset include `month` and `year` of the observation, `president`, indicating whether the President was a Democrat or a Republican on the associated date, `close`, indicating the closing values of the S&P stock index on the associated date, and `unemployment`, indicating the percent of unemployment on the associated date.

## Problem 3

```{r import_tidy_babynames}
baby_names_df = read_csv("Popular_Baby_Names.csv") %>%
                janitor::clean_names() %>%
                mutate(gender = str_to_lower(gender)) %>%
                mutate(childs_first_name = str_to_lower(childs_first_name)) %>%
                mutate(ethnicity = ifelse(grepl("ASIAN AND PACI", ethnicity), "ASIAN AND PACIFIC ISLANDER",
                                          ifelse(grepl("BLACK NON HISP", ethnicity), "BLACK NON HISPANIC",
                                                 ifelse(grepl("WHITE NON HISP", ethnicity), "WHITE NON HISPANIC",
                                                              "HISPANIC")))) %>%
                mutate(ethnicity = str_to_lower(ethnicity)) %>%
                distinct()
                                          
view(baby_names_df)
```

Produce a table showing the rank in popularity of teh name "Olivia" as a female baby name over time
```{r female_olivia}
olivia_df =
  filter(baby_names_df, childs_first_name == "olivia", gender == "female")

olivia_df %>%
  subset(select = -c(gender, childs_first_name, count)) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>%
  knitr::kable(caption = "The rank in popularity of the name Olivia as a female baby name over time")

view(olivia_df)
```

Produce a similar table showing the most popular name among male children over time.
```{r popular_male_name}
male_name = 
  filter(baby_names_df, gender == "male", rank == min(rank), count == max(count))

view(male_name)
```

Produce a scatterplot showing the number of children with a name against the rank in popularity of that name for male, white non-hispanic children born in 2016.
```{r scatterplot}
male_wnh_2016_df = 
  filter(baby_names_df, gender == "male", year_of_birth == "2016", ethnicity == "white non hispanic") %>%
  subset(select = -c(year_of_birth, ethnicity, gender))

male_wnh_2016_df %>%
  ggplot(aes(x = rank, y = count)) + geom_point()

view(male_wnh_2016)
```

